Comparison with Christensen paper

### 2019-01-08 paper-comparison/compare-christensen.Rmd
table summarizing cpt outcomes

#### run LDATS LDA with LDATS changepoint
```{r, include = F}
# remove paper LDA (for memory)
rm(ldamodel)
```

```{r LDATS LDA + LDATS cpt, eval = F}
# Run changepoint

ldats_ldats <- LDATS::TS_on_LDA(LDA_models = ldats_lda_selected,
                                document_covariate_table = rodents[[2]],
                                formulas =c(~ sin_year + cos_year),
                                nchangepoints = c(2, 3),
                                weights = NULL)

save(ldats_ldats, file = 'model-stash/ldats_ldats.Rds')

# Select best changepoint

ldats_ldats_selected <- LDATS::select_TS(TS_models = ldats_ldats)

save(ldats_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/ldats_ldats_selected.Rds')

@@

```{r, eval = F}
rm(ldats_ldats)
rm(ldats_ldats_selected)
```

#### run LDATS LDA with paper changepoint

```{r ldats paper, eval = F}
source('christensen-ecology-files/changepointmodel.r')
# set up parameters for model
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
  year_continuous = year_continuous,
  sin_year = sin(year_continuous * 2 * pi),
  cos_year = cos(year_continuous * 2 * pi)
)

# run models with 1, 2, 3, 4, 5 changepoints
cp_results_rodent = changepoint_model(ldats_lda_selected[[1]], x, 1, weights = rep(1,length(year_continuous)))

save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda1.Rds')
rm(cp_results_rodent)

cp_results_rodent2 = changepoint_model(ldats_lda_selected[[1]], x, 2, weights = rep(1,length(year_continuous)))

save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda2.Rds')
rm(cp_results_rodent2)

cp_results_rodent3 = changepoint_model(ldats_lda_selected[[1]], x, 3, weights = rep(1,length(year_continuous)))
save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda3.Rds')
rm(cp_results_rodent3)

cp_results_rodent4 = changepoint_model(ldats_lda_selected[[1]], x, 4, weights = rep(1,length(year_continuous)))
save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda4.Rds')
rm(cp_results_rodent4)

cp_results_rodent5 = changepoint_model(ldats_lda_selected[[1]], x, 5, weights = rep(1,length(year_continuous)))
save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda5.Rds')
rm(cp_results_rodent5)

cp_results_rodent6 = changepoint_model(ldats_lda_selected[[1]], x, 6, weights = rep(1,length(year_continuous)))
save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_ldats_lda6.Rds')
rm(cp_results_rodent6)

rm(ldats_lda_selected)
```

#### run paper LDA with paper changepoint
```{r paper paper, eval = F}
source('christensen-ecology-files/changepointmodel.r')
# set up parameters for model
year_continuous = 1970 + as.integer(julian(ch_dates)) / 365.25
x = data.frame(
year_continuous = year_continuous,
sin_year = sin(year_continuous * 2 * pi),
cos_year = cos(year_continuous * 2 * pi)
)
# run models with 1, 2, 3, 4, 5 changepoints
cp_results_rodent = changepoint_model(ldamodel, x, 1, weights = rep(1,length(year_continuous)))
save(cp_results_rodent, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda1.Rds')
rm(cp_results_rodent)

cp_results_rodent2 = changepoint_model(ldamodel, x, 2, weights = rep(1,length(year_continuous)))
save(cp_results_rodent2, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda2.Rds')
rm(cp_results_rodent2)

cp_results_rodent3 = changepoint_model(ldamodel, x, 3, weights = rep(1,length(year_continuous)))
save(cp_results_rodent3, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda3.Rds')
rm(cp_results_rodent3)

cp_results_rodent4 = changepoint_model(ldamodel, x, 4, weights = rep(1,length(year_continuous)))
save(cp_results_rodent4, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda4.Rds')
rm(cp_results_rodent4)

cp_results_rodent5 = changepoint_model(ldamodel, x, 5, weights = rep(1,length(year_continuous)))
save(cp_results_rodent5, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda5.Rds')
rm(cp_results_rodent5)

cp_results_rodent6 = changepoint_model(ldamodel, x, 6, weights = rep(1,length(year_continuous)))
save(cp_results_rodent6, file = '/Users/renatadiaz/Documents/model-stash/paper_cpt_paper_lda6.Rds')
rm(cp_results_rodent6)
```

#### run paper LDA with LDATS changepoint
```{r paper ldats, eval = F}
# Run changepoint

paper_ldats <- LDATS::TS_on_LDA(LDA_models = ldamodel,
                                document_covariate_table = rodents[[2]],
                                formulas =c(~ sin_year + cos_year),
                                nchangepoints = c(2, 3, 4, 5, 6),
                                weights = NULL)

save(paper_ldats, file = 'model-stash/paper_ldats.Rds')

# Select best changepoint

paper_ldats_selected <- LDATS::select_TS(TS_models = paper_ldats)

save(paper_ldats_selected, file = '/Users/renatadiaz/Documents/model-stash/paper_ldats_selected.Rds')

>>>>

For narratives running all these models, see `ldats_changepoint.md` and `paper_changepoint.md`.

All the models found 4 changepoints (although the paper model actually finds lower deviances for 6 changepoints, 6 was outside the range considered by the paper/anticipated by LDATS, so I'm sticking with 4 for now)

### Compare model outcomes

```{r read model results}
ldats_ldats <- read.csv('ldats_ldats_dates.csv')
ldats_paper <- read.csv('ldats_paper_dates.csv')
paper_ldats <- read.csv('paper_ldats_dates.csv')
paper_paper <- read.csv('paper_paper_dates.csv')

estimates <- c(1:4)
estimates <- as.data.frame(estimates)

# Naming convention: ldamethod_cptmethod
# so ldats_paper is the ldats lda and paper cpt, etc
estimates$ldats_ldats <- ldats_ldats$newmoondate
estimates$ldats_paper <- ldats_paper$cpt_dates
estimates$paper_ldats <- paper_ldats$newmoondate
estimates$paper_paper <- paper_paper$cpt_dates

estimates

@@
